From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-04 WOLFHPC15/article/component.tex
wip Julien

%----------------------------------------
%\subsection{Overview}
%----------------------------------------
This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
Before those details the section gives an overview of component models, and introduce the Low Level Components (\llc), the model used in this paper.

@@

\llc is a minimalist \texttt{C++} based HPC-oriented component model
where a component extends the concept of class.
The services offered by the components are specified trough $provide$ ports,
those used either by $use$ ports for a single service instance,
or $use-multiple$ ports for multiple service instances.
Services are specified as \texttt{C++} interfaces.
\llc also offers $MPI$ ports that enable components to share MPI communicators.
Components can also have attribute ports to be configured.
%
As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
represented by a white circle, a $use$
port with a black circle, a $use-multiple$ port by a black circle with
a white $m$ in it. MPI port are
connected with a black rectangle.

\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}

A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
Such an assembly is described in LAD, an XML dialect~\cite{l2c}.

@@

@@

>>>>

%----------------------------------------
%\subsection{Overview}
%----------------------------------------
This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
Before those details, the section gives an overview of component models, and especially the Low Level Components (\llc) used in this paper.

@@

\llc is a minimalist \texttt{C++} based HPC-oriented component model
where a component extends the concept of class.
The services offered by the components are specified trough $provide$ ports,
those used either by $use$ ports for a single service instance,
or $use-multiple$ ports for multiple service instances.
Services are specified as \texttt{C++} interfaces.
\llc also offers $MPI$ ports that enable components to share MPI communicators.
Components can also have attribute ports to be configured.
%
As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
represented by a white circle, a $use$
port with a black circle, a $use-multiple$ port by a black circle with
a white $m$ in it. MPI port are
connected with a black rectangle.
A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
Such an assembly is described in LAD, an XML dialect~\cite{l2c}.
\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}

@@

\todo[inline]{JB: debut WIP}

The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support. A general view of this assembly is proposed in Figure~\ref{fig:mscac:assembly}; it is explained in details in Section~\ref{sect:component}. As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.

The final back-end component assembly is a ready-to-fill parallel structure where the user has to write computation codes described in the MSL description. To write those computations, interfaces of the choosen implementation have to be used. For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}. A work in progress replaces this DDS component by another one making use of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (D) at (0,0) {$Driver$};
   \node[provide] (Dp) at (-1,0) {};
   \node (Ds) at (-1.5,0) {start};
   \node[use,right=1.5cm of D] (Du1) {};
   \node[use,below=0.8cm of D] (Du2) {};

   \node[provide,below=0.15 of Du2] (DAp) {};
   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
   \node[use,right=1.7cm of DA] (DAu1) {$m$};
   \node[use,below=0.8cm of DA] (DAu2) {};

   \node[provide,below=0.15 of DAu2] (Tp) {};
   \node[component,below=0.8cm of Tp] (T) {$Time$};
   \node[use,right=1cm of T] (Tu) {};
   \node[left=0.8cm of T] (tt) {$T$};

   \node[provide,right=0.15 of Tu] (Cp) {};
   \node[component,right=2cm of Cp] (C) {$Computations$};
   \node[use,above=0.8cm of C] (Cu) {};
   \node[right=0.2cm of Cu] (star) {$*$};
   \node[right=1.5cm of C] (gamma) {$\Gamma$};

   \node[provide,right=0.2 of DAu1] (Datap1) {};
   \node[provide,above=0.15 of Cu] (Datap2) {};
   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};

   \node[provide,right=0.15 of Du1] (DDSp1) {};
   \node[provide,above=0.15 of Datau] (DDSp2) {};
   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
    (DDSp1) edge node {} (DDS)
    (DA) edge node {} (DAu1)
           edge node {} (DAu2)
    (DAp) edge node {} (DA)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\caption{MSCAC Component Assembly Back-End Overview.}
\label{fig:mscac:assembly}
\end{center}
\end{figure}

@@

\todo[inline]{JB: fin WIP}

### 2015-08-04 WOLFHPC15/article/concept.tex
wip Julien

A multi-stencil program in this work is defined as a sextuplet:
\begin{equation} 
\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma),
\label{eq:msp}
\end{equation}
where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used into computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations. The set of computations $\Gamma$ is an ordered list of stencil kernels and local computations.

>>>>

A multi-stencil program in this work is defined as a sextuplet:
\begin{equation} 
\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma),
\label{eq:msp}
\end{equation}
where $T$ is the set of time iteration to run the simulation, $\mathcal{M}$ is the mesh of the simulation, $\mathcal{E}$ is the set of mesh entities, $\mathcal{D}$ is the set of computation domains used for computations, $\Delta$ is the set of data of the simulation, each one mapped onto a mesh entity, and finally $\Gamma$ is the set of computations. The set of computations $\Gamma$ is an ordered list of stencil kernels and local computations.

### 2015-08-04 WOLFHPC15/article/ms_msc.tex
wip Julien

As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations. Six main sections are required in a MSL description that match the six tuples of the formal definition of a multi-stencil program: \textbf{1.}\,the mesh description, \textbf{2.}\,the mesh entities description, \textbf{3.}\,the computation domains description and their dependencies, \textbf{5.}\,the data description, \textbf{5.}\,the time loop description, and finally \textbf{6.} the computations description.
Figure~\ref{fig:grammar} shows the grammar of MSL. Lines 1 to 9 define what is a MSL program with the different parts mentioned above. The remaining of this section describes those different parts of a MSL program and presents an example.
We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers. Only \texttt{num} is an integer terminal.

\subsubsection*{Mesh and Mesh Entities}
As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier. Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}). Figure~\ref{fig:mslex} presents an example defining a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}.

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain can be seen as a subpart of entities of a mesh which will be used during one or more computation. For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}). By default two computation domains always intersect except when an independence relation is mentioned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
In the lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they represent the whole domain.
On lines 7 and 8 two other subparts of the mesh entity \texttt{edgex} are defined, called \texttt{part1edgex} and \texttt{part2edgex}; they represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified since they do not intersect.

\subsubsection*{Data and Time}
A data is a quantity to simulate. It is mapped on mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}. The \texttt{time} section simply indicates a number of iterations to perform in the simulation. In the example of Figure~\ref{fig:mslex} ten data identifiers are defined and 500 iterations (lines 11-22). %An extension could be to support a convergence criterion.

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The specification starts by the identifier of the data to compute ($w$ in the formal description) with the computation domain between brackets ($d$ in the formal definition).
Then, after the equal sign is specified the kernel identifier.
Finally, between the parenthesis is the list of identifiers of data read by the computation with their stencil shape between brackets ($R$ in the formal definition).
If the computation is local, no brackets appear ($R_l$ in the formal definition). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

@@

%----------------------------------------
\subsection{MSC Parallelization}
%----------------------------------------
The MSC parallelization is a subpart of the overall compiler MSCAC.
It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
Data parallelism is handled in the first two step, and task parallelism in the following three steps. This section gives an overview of the five steps of the MSC transformation.
The detailed formalism of these steps and the associated algorithms are described in a research report~\cite{????}.

@@

\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
This does however require additional synchronizations between processors.
For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
A synchronization is needed each time a data read by a stencil computation has been written by a previous computation. In such a case, a computation is added before the stencil computation. This \emph{synchronization computation} reads the data to synchronize, and write the same data. The computation domain of such a synchronization computation is the mesh entity on which the data is declared. As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ read the data $b$ which has been written by the computation $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed to the sublist $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$. The new computation $sync_1$ read and write $b$ and is applied on the computation domain of $c_1$. As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$. The same transformation is performed for all stencils: $c_4$ and $c_8$. Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.

\subsubsection*{The Dependency Graph: $\Gamma_{dep}$}
From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built. A dependency exists between two computations (including synchronizations) if and only if a data read has been written by a previous computation in $\Gamma_{sync}$, and if the computation domains intersect. In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect. Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them. The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{DAG}). For example, the dependency DAG $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is presented in Figure~\ref{fig:hyb}.

@@

>>>>

As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations. Six main sections are required in a MSL description that match the six elements of the formal definition; they describe: \textbf{1.}\,the mesh, \textbf{2.}\,the mesh entities, \textbf{3.}\,the computation domains with their inter-dependencies, \textbf{4.}\,the data, \textbf{5.}\,the time iterations, and finally \textbf{6.} the computations.
Figure~\ref{fig:grammar} shows the grammar of MSL. Lines 1 to 9 define what is a MSL program with the different parts mentioned above. The remaining of this section describes those different parts of a MSL program and presents an example.
We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers. Only \texttt{num} is an integer terminal.

\subsubsection*{Mesh and Mesh Entities}
As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
Kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
Figure~\ref{fig:mslex} presents an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kind of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain is a subset of a given kind of entities from a mesh used during one or more computations.
For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
By default two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
On lines 7 and 8 two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.

\subsubsection*{Data and Time}
A data element is a quantity to simulate. It is mapped on a given kind of mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}. The \texttt{time} section simply indicates a number of iterations to perform in the simulation. In the example of Figure~\ref{fig:mslex} ten data elements are defined and 500 iterations (lines 11-22). %An extension could be to support a convergence criterion.

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The specification starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
Then, after the equal sign, the kernel identifier ($exp$) is specified.
Finally, between the parenthesis is the list of identifiers of the data elements read by the computation with their stencil shape between brackets ($R$).
If the computation is local, no brackets appear ($R_l$). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

@@

%----------------------------------------
\subsection{MSC Parallelization}
%----------------------------------------
The MSC parallelization is a subpart of the overall compiler MSCAC.
It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.

@@

\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
This does however require additional synchronizations between processors.
For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
A synchronization is needed each time a data read by a stencil computation has been written by a previous computation. In such a case, a computation is added before the stencil computation. This \emph{synchronization computation} reads the data to synchronize, and write the same data. The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared. As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$. The new computation $sync_1$ reads and writes $b$ and is applied on the computation domain of $c_1$. As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$. The same transformation is performed for all stencils: $c_4$ and $c_8$. Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
\todo[inline]{JB: "is applied on the computation domain of $c_1$" vraiment, c'est pas ce que tu dis juste avant ...}

\subsubsection*{The Dependency Graph: $\Gamma_{dep}$}
From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built.
A dependency exists from computation $a$ to $b$ (including synchronizations) if and only if a data element read by $b$ has been written by $a$ in $\Gamma_{sync}$ with intersecting domains.
%In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them. The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{DAG}). For example, the dependency DAG $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is presented in Figure~\ref{fig:hyb}.

@@

\todo[inline]{JB: dans le DAG il y a une dependance de c2 a sync4, le graphe serie parallele les execute en parallele ...}

### 2015-08-04 WOLFHPC15/article/mscac.tex
5.4,5.5
wip Julien

The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
Various models could be used as back-end, such as OpenMP~\cite{660313} for example, however this paper adopts a novel approach where the dump targets a component-based runtime.
This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
%where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for

\todo[inline]{JB: je déplacerais les 2 derniers paragraphes et la figure dans la sec.~\ref{sect:component}}

The back-end of the compiler builds a component assembly, that can be divided in two parts: the components implementing the actual computation and the static schedule, and the components needed to manage it, \ie the runtime support. A general view of this assembly is proposed in Figure~\ref{fig:mscac:assembly}; it is explained in details in Section~\ref{sect:component}. As illustrated in the figure, each component is linked to one or more tuple of the multi-stencil program definition: \emph{DDS} component represents $\mathcal{M}$ and $\mathcal{E}$; \emph{Data} component represents $\Delta$ and $\mathcal{D}$; \emph{Time} component represents $T$, and \emph{Computations} component (which is generated by MSC) represents $\Gamma$.

The final back-end component assembly is a ready-to-fill parallel structure where the user has to write computation codes described in the MSL description. To write those computations, interfaces of the choosen implementation have to be used. For example, the back-end evaluated in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}. A work in progress replaces this DDS component by another one making use of Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985}.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (D) at (0,0) {$Driver$};
   \node[provide] (Dp) at (-1,0) {};
   \node (Ds) at (-1.5,0) {start};
   \node[use,right=1.5cm of D] (Du1) {};
   \node[use,below=0.8cm of D] (Du2) {};

   \node[provide,below=0.15 of Du2] (DAp) {};
   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
   \node[use,right=1.7cm of DA] (DAu1) {$m$};
   \node[use,below=0.8cm of DA] (DAu2) {};

   \node[provide,below=0.15 of DAu2] (Tp) {};
   \node[component,below=0.8cm of Tp] (T) {$Time$};
   \node[use,right=1cm of T] (Tu) {};
   \node[left=0.8cm of T] (tt) {$T$};

   \node[provide,right=0.15 of Tu] (Cp) {};
   \node[component,right=2cm of Cp] (C) {$Computations$};
   \node[use,above=0.8cm of C] (Cu) {};
   \node[right=0.2cm of Cu] (star) {$*$};
   \node[right=1.5cm of C] (gamma) {$\Gamma$};

   \node[provide,right=0.2 of DAu1] (Datap1) {};
   \node[provide,above=0.15 of Cu] (Datap2) {};
   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};

   \node[provide,right=0.15 of Du1] (DDSp1) {};
   \node[provide,above=0.15 of Datau] (DDSp2) {};
   \node[component,above=0.8cm of DDSp2] (DDS) {$DDS$};
   \node[right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
    (DDSp1) edge node {} (DDS)
    (DA) edge node {} (DAu1)
           edge node {} (DAu2)
    (DAp) edge node {} (DA)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\caption{MSCAC Component Assembly Back-End Overview.}
\label{fig:mscac:assembly}
\end{center}
\end{figure}

>>>>

The final phase of compilation, the dump, uses the static scheduling of the tasks, created from $\Gamma$, and additional informations on $T$, $\mathcal{M}$, $\Delta$, $\mathcal{E}$ and $\mathcal{D}$.
It generates a program in the back-end language which embeds the parallel pattern or skeleton of the overall simulation.
Various models could be used as back-end, such as OpenMP~\cite{660313} for example, however this paper adopts a novel approach where the dump targets a component-based runtime.
This approach makes it possible to leverage good properties of software component models such as increased code re-use, eased separation of concerns, and increased maintainability and productivity.
% JB: la phrase en dessous est pas tres heureuse ...
Those properties counter some known limitations of DSLs. % JB: was "parallel applications".
%where a component represents a clear and independent functionality of the application, have first been proposed in software engineering and have proved many times good properties for

Thanks to the software components based approach, the generated code is made of independent parts:
A first set of components encodes the static scheduling provided by MSC.
The other parts loosely match the elements of the multi-stencil program definition, these include the components implementing: \textbf{a)} the kernels ($\Gamma$), \textbf{b)} the data structures ($\mathcal{M}$ and $\mathcal{E}$), \textbf{c)} their parallelization ($\Delta$ and $\mathcal{D}$), and \textbf{d)} the time loop ($T$).
This enables to change the technology used for each part while limiting the impact on the other parts as further discussed in Section~\ref{sect:component}.
